I don't know if projection is categorical. \\ Did \citealt{mandelkern-etal2020} discover that it is?

### 2023-12-07 output/naturalness-paper/paper.tex
updated paper
minor paper edits

One hypothesis is that naturalness ratings in explicit ignorance context are sensitive not just to whether the presupposition is defeasible (which is not the case for {\em too, also} and {\em again}) but also to at-issueness. On this hypothesis, the explicit ignorance statement (e.g., {\em I don't know whether Julian dances salsa}) does not merely convey the speaker's ignorance about the content $c$ to be investigated, but also identifies that the speaker takes the question under discussion (QUD) to be whether $c$ is true (?$c$ e.g., {\em Does Julian dances salsa?}). The speaker's immediately following interrogative utterance is interpreted with respect to the QUD ?$c$. The naturalness of the interrogative is then modulated by whether the QUD ?$c$  and the interpretation of the interrogative constitute a felicitous strategy of inquiry (\citealt[32f.]{roberts12}), that is,  by whether the interrogative can be interpreted as inquiring about $c$. As shown in \citealt{tbd-variability}, projective content differs in whether it can be at-issue. The CCs of emotive predicates, including {\em be annoyed}, are particularly resistant to being interpreted as at-issue. Thus, an interrogative with {\em be annoyed} (e.g., {\em Is Cole annoyed that Julian dances salsa?}) is unlikely to be interpreted as congruent with a QUD about the CC (e.g., {\em Does Julian dance salsa?}). This incongruence may be why the CC of {\em be annoyed} received quite low naturalness ratings in the explicit ignorance context.

@@

In sum: The linking function assumed in \citealt{mandelkern-etal2020} only considers that naturalness ratings in explicit ignorance contexts are sensitive to whether content is coded as a presupposition. This linking function may be taken to be supported for expressions like {\em continue, too, also} and {\em again}, which received low naturalness ratings in explicit ignorance contexts and which have been analyzed as presupposition triggers. The results of our experiment for the CCs of {\em be annoyed} and {\em know}, however, suggest that naturalness ratings in explicit ignorance contexts may be sensitive to whether the CC can be at-issue and to the (possibly implicit) prosody of the utterances. Consequently, one cannot draw conclusions from the mean naturalness rating that an expression received in the explicit ignorance context to its status as a presupposition trigger, contrary to what is suggested by the linking function assumed in \citealt{mandelkern-etal2020}.

@@

\subsection{Analyzing projection variability}

@@

As mentioned in \S\ref{s1}, \citealt{tbd-variability} proposed, based on the results of several experiments that used an inference rating measure, that projection is gradient.  \citealt{mandelkern-etal2020}, on the other hand, claimed that their naturalness rating measure categorically distinguishes presuppositions and nonpresuppositions.  However, as discussed above, neither the results presented in \citealt{mandelkern-etal2020} nor the results of our experiment empirically support the claim that projection is categorical and, instead, further supports the proposal that it is gradient. In other words, the empirical evidence that is available to date, and which is now based on both inference and naturalness rating measures, supports the proposal that by-content projection variability is gradient, not categorical.

This result gives rise to the question of how to analyze the observed projection variability. The analyses in \citealt{heim83} and \citealt{vds92} assumes that expressions are either presupposition triggers or not, and they rely on local accommodation to account for examples in which presuppositions do not project.  As discussed in detail in \citealt{tbd-variability}, it is not clear that local accommodation can capture the observed by-content variation in a principled way, as analyses would need to assume that the CC of {\em discover} is more amenable to local accommodation than the CC of {\em know}. Another problem such analyses face is that the CC of nonfactive predicates is also projective, sometimes even more so than the CC of factive predicates. Thus, as also discussed in \citealt{degen-tonhauser-language}, it is not clear which expressions would be analyzed as triggering projection, and which ones not.

Another set of projection analyses was developed to capture the projection of content associated with soft triggers (e.g., \citealt{abusch02,abusch10, abrusan2011,abrusan2016,romoli2015,simons-etal2017}). While these analyses differ in detail, they were developed to capture that the projection inferences associated with soft triggers are more easily suspended (that is, less projective) than those of hard triggers. The availability of such analyses means a three-way distinction in projection can already be made: Content that is conventionally marked as a presupposition (like the prejacent of {\em too} and {\em also}) is expected to be more projective than the content of soft triggers (like the prejacent of {\em stop} or the CC of {\em discover}), which in turn is expected to be more projective than content contributed by other expressions (like the CC of {\em think}). While this is promising, contemporary projection analyses do not suffice to account for observed by-content projection variability between soft triggers (e.g., that the CC of {\em know} is more projective than that of {\em discover}, which in turn is more projective than that of {\em reveal}), or that contents associated with nontriggers, such as the CCs of nonfactive predicates, may be projective, too, albeit to varying degrees (see \citealt{tbd-variability} and \citealt{degen-tonhauser-language} for discussion).

This situation suggests a promising avenue for future research that picks up a theme from \S\ref{s31}, namely to continue investigating distinctions between projection inferences. Understanding the dimensions on which projection inferences differ (or not differ) will lead to a refined taxonomy of such inferences (for a preliminary taxonomy see \citealt{brst-lang11}). The differences between empirically-motivated classes of projection inferences can, in turn, be decisive in identifying inferences already captured by extant analyses and inferences for which novel analyses must be developed.

>>>>

One hypothesis is that naturalness ratings in explicit ignorance contexts are sensitive not just to whether the presupposition is defeasible (which is not the case for {\em too, also} and {\em again}) but also to at-issueness. On this hypothesis, the explicit ignorance statement (e.g., {\em I don't know whether Julian dances salsa}) does not merely convey the speaker's ignorance about the content $c$ to be investigated, but also identifies that the speaker takes the question under discussion (QUD) to be whether $c$ is true (?$c$ e.g., {\em Does Julian dances salsa?}). The speaker's immediately following interrogative utterance is interpreted with respect to the QUD ?$c$. The naturalness of the interrogative is then modulated by whether the QUD ?$c$  and the interpretation of the interrogative constitute a felicitous strategy of inquiry (\citealt[32f.]{roberts12}), that is,  by whether the interrogative can be interpreted as inquiring about $c$. As shown in \citealt{tbd-variability}, projective content differs in whether it can be at-issue. The CCs of emotive predicates, including {\em be annoyed}, are particularly resistant to being interpreted as at-issue. Thus, an interrogative with {\em be annoyed} (e.g., {\em Is Cole annoyed that Julian dances salsa?}) is unlikely to be interpreted as congruent with a QUD about the CC (e.g., {\em Does Julian dance salsa?}). This incongruence may be why the CC of {\em be annoyed} received quite low naturalness ratings in the explicit ignorance context.

@@

In sum: The linking function assumed in \citealt{mandelkern-etal2020} only considers that naturalness ratings in explicit ignorance contexts are sensitive to whether content is coded as a presupposition. This linking function may be taken to be supported for expressions like {\em continue, too, also} and {\em again}, which received low naturalness ratings in explicit ignorance contexts and which have been analyzed as presupposition triggers. The results of our experiment for the CCs of {\em be annoyed} and {\em know}, however, suggest that naturalness ratings in explicit ignorance contexts may be sensitive to whether the CC can be at-issue and to the (possibly implicit) prosody of the utterances. Consequently, one cannot draw conclusions from an expression's mean naturalness rating in an explicit ignorance context to its status as a presupposition trigger, contrary to what is suggested by the linking function assumed in \citealt{mandelkern-etal2020}.

@@

\subsection{Theoretical analyses of projection variability}

@@

As mentioned in \S\ref{s1}, \citealt{tbd-variability} proposed, based on the results of several experiments that used an inference rating measure, that projection is gradient.  \citealt{mandelkern-etal2020}, on the other hand, claimed that their naturalness rating measure categorically distinguishes presuppositions and nonpresuppositions.  However, as discussed above, neither the results presented in \citealt{mandelkern-etal2020} nor the results of our experiment empirically support the claim that projection is categorical and, instead, further supports the proposal that it is gradient. In other words, the empirical evidence that is available to date, and which is now based on both inference and naturalness rating measures, supports the proposal that by-content projection variability is gradient, not categorical. \jd{what does it mean for variability to be gradient? i thought the argument is that projection means form a gradient and there are different things that might mean theoretically. but i'm not sure what you have in mind when you say "by-content projection variability is gradient"}

This raises the question of how to analyze the observed projection variability. The analyses in \citealt{heim83} and \citealt{vds92} assume that expressions are either presupposition triggers or not, and they rely on local accommodation to account for examples in which presuppositions do not project.  As discussed in detail in \citealt{tbd-variability}, it is not clear that local accommodation can capture the observed by-content variation in a principled way, as analyses would need to assume that the CC of {\em discover} is more amenable to local accommodation than the CC of {\em know}. Another problem such analyses face is that the CC of nonfactive predicates is also projective, sometimes even more so than the CC of factive predicates. Thus, as also discussed in \citealt{degen-tonhauser-language}, it is not clear which expressions would be analyzed as triggering projection, and which ones not.

Another set of projection analyses was developed to capture the projection of content associated with soft triggers (e.g., \citealt{abusch02,abusch10, abrusan2011,abrusan2016,romoli2015,simons-etal2017}). While these analyses differ in detail, they were developed to capture that the projection inferences associated with soft triggers are more easily suspended than those of hard triggers. These analyses predict s a three-way distinction in projection: Content that is conventionally marked as a presupposition (like the prejacent of {\em too} and {\em also}) is expected to be more projective than the content of soft triggers (like the prejacent of {\em stop} or the CC of {\em discover}), which in turn is expected to be more projective than content contributed by other expressions (like the CC of {\em think}). While this is promising, contemporary projection analyses do not suffice to account for observed by-content projection variability between soft triggers (e.g., that the CC of {\em know} is more projective than that of {\em discover}, which in turn is more projective than that of {\em reveal}), or that contents associated with nontriggers, such as the CCs of nonfactive predicates, may be projective, too, albeit to varying degrees (see \citealt{tbd-variability} and \citealt{degen-tonhauser-language} for discussion).

This situation suggests a promising avenue for future research that picks up a theme from \S\ref{s31}, namely to continue investigating distinctions between projection inferences. Understanding the dimensions along which projection inferences do or don't differ will lead to a refined taxonomy of such inferences (for a preliminary taxonomy see \citealt{brst-lang11}). The differences between empirically-motivated classes of projection inferences can, in turn, be decisive in identifying inferences already captured by extant analyses and inferences for which novel analyses must be developed.
